---
title: "assessment-final-mt"
author: "My My Tran"
date: "2024-06-22"
output: pdf_document
---

# Introduction

Network meta-analysis (NMA) allows for the comparative relative efficacy and acceptability of interventions, such as galcanezumab and other treatments for chronic migraine, through randomized controlled trials (RCTs) [1]. These clinical trials form a network of observational evidence, allowing for both direct and indirect comparisons. Prior to this NMA, a feasibility assessment was completed, revealing minimal concerns regarding heterogeneity across studies, implying that an NMA is appropriate. This NMA evaluates the reduction in monthly migraine days for galcanezumab in comparison to other treatments, including key comparators such as Botox A, eptinezumab, erenumab, and placebo.

# Data and Methods

This dataset is based on multiple 12-week observational studies of clinical trials conducted on participants, on their reduction in monthly migraine days from baseline to Week 12. It is comprised of the relative mean difference versus placebo in the reduction of monthly migraine days (y), the standard error in the mean difference (se), and the number of treatment arms in the study (na).

We transform the data from long format into a pairwise format, separating the treatment types into treatment 1 and treatment 2. These studies provide direct comparisons against the placebo. Using placebo as the common comparator, we can estimate the relative efficacy between two interventions indirectly.

When constructing the network meta-analysis, comparisons with missing treatment effect point estimates (TE), seTE, zero TE values are not considered in the analysis. Therefore, two studies, Jones (1995), Gaudi (2001), are excluded, which loses pieces of evidence for the treatments galcanezumab and erenumab, respectively. Additionally, there is one three-arm study in this dataset. Since each comparison in the three-arm study should have its own effect size and standard error, we provide data for each pairwise comparison [2]. This includes (1) galcanezumab vs placebo, (2) galcanezumab vs eptinezumab, and (3) placebo vs eptinezumab. Given the MD and SE between galcanezumab and placebo, it is estimated that the MD between galcanezumab and eptinezumab is 0.70 and the SE is 0.7401. 

Therefore, the NMA includes data from 11 RCTS, comprising 13 pairwise comparisons among 5 treatments (Botox A, eptinezumab, erenumab, galcanezumab, and placebo). 

The method used was a Frequentist random effects model to account for heterogeneity or variability across studies, even if minimal. We are interested in the random effects model as it allows for the possibility that true effect sizes vary across studies. The studies can differ in patient populations, methodologies, and other treatment-influencing factors [3]. Assumptions of the Frequentist method include transitivity, which are different treatments that are similar in severity of chronic migraines, treatment dose or study quality, and congruence and consistency [4].

To conduct the NMA using the Frequentist approach, we use the netmeta package in R [5]. The chosen reference treatment, or the baseline against which all other treatments are compared is the placebo. This allows us to see how galcanezumab and other treatments compare to this same baseline [6].

# Results

```{r}
#| echo: false
#| warning: false
library(tidyverse)
library(readxl)
library(netmeta)
chronic_migraine_dataset <- read_excel("Chronic_Migraine_Dataset.xlsx")
transformed_migraine_dataset <- chronic_migraine_dataset |>
  group_by(study, year) |>
  mutate(trt1name = ifelse(trt != "Placebo", trt, NA),
         trt2name = ifelse(trt == "Placebo", "Placebo", NA)) |>
  summarise_all(list(~ first(na.omit(.))))
transformed_migraine_dataset$y <- as.numeric(transformed_migraine_dataset$y)
transformed_migraine_dataset$se <- as.numeric(transformed_migraine_dataset$se)
transformed_migraine_dataset <- transformed_migraine_dataset |>
  drop_na(se) |>
  select(-trt)
crisp_arm_g <- data.frame(study = "Crisp", year = "2012", y = 0.70, se = 0.7401, na = 3, trt1name = "Galcanezumab", trt2name = "Eptinezumab")

crisp_arm_e <- data.frame(study = "Crisp", year = "2012", y = -4.90, se = 0.5669, na = 3, trt1name = "Eptinezumab", trt2name = "Placebo")

transformed_migraine_dataset <- transformed_migraine_dataset |>
  bind_rows(crisp_arm_g, crisp_arm_e) |>
  arrange(year)
```

```{r}
#| echo: false
#| warning: false
nma <- netmeta(TE = y, seTE = se, treat1 = trt1name, treat2 = trt2name, studlab = study, data = transformed_migraine_dataset, sm = "MD", ref = "Placebo", common = FALSE, small.values = "desirable")
```

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: This is a graphical visualization of the network meta-analysis, which compares all treatments within the network. The nodes represent interventions (treatment types), and edges are direct comparisons, weighted by the number of studies. Erenumab and galcanezumab have the most studies in this dataset. 

netgraph(nma, plastic = FALSE, points = TRUE, col = 'darkblue', thickness = "number.of.studies", lwd = 2.7, cex.points = 4, offset = 0.05, scale = 1.1, col.points = 'red', seq = 1)
```
```{r}
#| echo: false
#| warning: false
#| label: table-1
#| fig-cap: The table summarizes the estimated mean differences in migraine days per month for each treatment compared to placebo.

print(nma)
```

The four treatments (Botox A, Eptinezumab, Erenumab, and Galcanezumab reveal significance in reducing the monthly days of a migraine days (alpha = 0.05), with Eptinezumab and Erenumab having the largest effects (p < 0.001). Additionally, heterogeneity was detected (p < 0.0001), suggesting there is variability in treatment effects across studies.

```{r}
#| echo: false
#| warning: false
#| label: table-2
#| fig-cap: There are 121,577 observations in the original dataset along with summary statistics (minimum, maximum, median, mean, and quantiles) for our variables of interest. Our predictor variables are race (categorical), deyo (median of 1 and mean of 1.023, indicating more patients with mild severity), and our rescaled variable age_cent has a mean of 0. Length of stay (LOS_binary) has a mean of 0.022.

forest(nma)
```

```{r}
#| echo: false
#| warning: false
#| label: table-3
#| fig-cap: There are 121,577 observations in the original dataset along with summary statistics (minimum, maximum, median, mean, and quantiles) for our variables of interest. Our predictor variables are race (categorical), deyo (median of 1 and mean of 1.023, indicating more patients with mild severity), and our rescaled variable age_cent has a mean of 0. Length of stay (LOS_binary) has a mean of 0.022.

set.seed(76)
rankogram(nma) # probabilities that any of our 5 treatments are the best treatment 

pg <- rankogram(nma, cumulative = TRUE)

plot(pg)

netrank(pg) # average proportion that any of our treatments is better than other interventions
```

# Concerns with regard to the dataset. 

# Interpretation of any treatment ranking statistics 










# Appendix

```{r}
library(tidyverse)
library(readxl)
library(netmeta)
```


```{r}
chronic_migraine_dataset <- read_excel("Chronic_Migraine_Dataset.xlsx")
```


```{r}
# transform dataset to separate treatment type and placebo
transformed_migraine_dataset <- chronic_migraine_dataset |>
  group_by(study, year) |>
  mutate(trt1name = ifelse(trt != "Placebo", trt, NA),
         trt2name = ifelse(trt == "Placebo", "Placebo", NA)) |>
  summarise_all(list(~ first(na.omit(.))))
```

```{r}
transformed_migraine_dataset$y <- as.numeric(transformed_migraine_dataset$y)
transformed_migraine_dataset$se <- as.numeric(transformed_migraine_dataset$se)
```

```{r}
# remove TEs or seTEs that are NA
transformed_migraine_dataset <- transformed_migraine_dataset |>
  drop_na(se) |>
  select(-trt)
```

```{r}

crisp_arm_g <- data.frame(study = "Crisp", year = "2012", y = 0.70, se = 0.7401, na = 3, trt1name = "Galcanezumab", trt2name = "Eptinezumab")

crisp_arm_e <- data.frame(study = "Crisp", year = "2012", y = -4.90, se = 0.5669, na = 3, trt1name = "Eptinezumab", trt2name = "Placebo")

transformed_migraine_dataset <- transformed_migraine_dataset |>
  bind_rows(crisp_arm_g, crisp_arm_e) |>
  arrange(year)
```

```{r}
# create net meta-analysis
nma <- netmeta(TE = y, seTE = se, treat1 = trt1name, treat2 = trt2name, studlab = study, data = transformed_migraine_dataset, sm = "MD", ref = "Placebo", common = FALSE, small.values = "desirable")
print(nma)
```

The treatments reduce the monthly days of a migraine days. All treatments are significant, with Eptinezumab and Erenumab having the largest effects (p < 0.001). 

```{r}
netgraph(nma, plastic = FALSE, points = TRUE, col = 'darkblue', thickness = "number.of.studies", lwd = 2.7, cex.points = 4, offset = 0.05, scale = 1.1, col.points = 'red', seq = 1)
```

```{r}
# Basic network effects forest plot
forest(nma)
```

Standardized mean differences
Larger confidence intervals, less precision

All of the studies do not cross the line of no effect, revealing significance

Represents the effect size, such as mean difference (MD), odds ratio (OR), risk ratio (RR), etc.
Usually includes a line at 0 (for MD) or 1 (for OR/RR) indicating no effect or no difference between treatments.
A vertical line at 0 (for MD) or 1 (for OR/RR) represents the line of no effect.
Each horizontal line represents the confidence interval (CI) of the effect size for a particular comparison.
The square or diamond at the center of each line represents the point estimate of the effect size.
For MD, a point estimate to the left of 0 indicates that the treatment is better than the comparator (e.g., less pain), 
Narrow intervals indicate more precise estimates, while wider intervals indicate less precision.

Eptinezumab shows the most precision shows a significant reduction in migraines, with a narrower CI indicating more precise estimates.

```{r}
set.seed(76)
rankogram(nma) # probabilities that any of our 5 treatments are the best treatment 

pg <- rankogram(nma, cumulative = TRUE)

plot(pg)

netrank(pg) # average proportion that any of our treatments is better than other interventions

```



It displays the probabilities (based on simulations) of each treatment being ranked from 1 to 5. 
The output you've provided shows the probabilities for each treatment (Botox A, Eptinezumab, Erenumab, Galcanezumab, Placebo) being ranked from 1st to 5th based on 1000 simulations.

For Botox A, in 10.5% of simulations, it was ranked 1st, in 16.2% it was ranked 2nd, and so on.
For Placebo, in 99.1% of simulations, it was ranked 5th.

The rankogram is based on results from a random effects model assumption.
This model allows for variability in treatment effects across studies, which is reflected in the distribution of ranks seen in the rankogram.

When heterogeneity is minimal, a fixed effects model might be appropriate. 
When heterogeneity is minimal, the assumption that all studies are estimating the same underlying effect size is more likely to hold true.
This simplifies the model as less variation needs to be accounted for, making the frequentist approach both efficient and effective.
If there is some minor heterogeneity, a random effects model can still be used to account for this small degree of variability.

Eptinezumab and Erenumab are likely to be the most effective treatments, followed by Botox A and Galcanezumab.

# References